---
title: "Rational Decisions"
author: "Your Name Here"
date: "The Date Here"
output: 
  html_document:
    fig_height: 3
    fig_width: 5
  pdf_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---

```{r include=FALSE}
require(knitr)
require(mosaic)
require(tigerstats)
```

# Introduction

<Do not modify the text in this section:  it sets the assignment for you.>

Do people make decisions rationally, based only upon information that is relevant to the issue they must decide, or is their decision affected by factors unrelated to finance?

This question was studied in a survey of 267 MAT 111 students at Georgetown College in the fall semester of 2022.  Each student was given a survey form that included a question about attendance at a rock concert, but not all forms were the same.  In fact, there were two types of form.

```{r eval=FALSE,include=FALSE}
#for the benedfit of a reader of the R Markdown source file:
data(attitudes)
View(attitudes)
help(attitudes)
```

Some of the subjects were looking at forms where the question was stated as follows:

  >"You have purchased a $30 ticket to see a rock concert in Rupp Arena. When you arrive at the Arena on the night of the performance, you find that you have lost the ticket. You have no receipt, so it will not be possible to see the concert unless you purchase another ticket. Would you purchase another ticket?"

The rest of the subjects were looking at forms where the question was stated as follows:

  >"You plan to see a rock concert in Rupp Arena. Tickets for the performance are $30. When you arrive at the Arena on the night of the performance, you find that you have lost two bills from your purse or wallet: a ten and a twenty. Would you buy the ticket anyway?"

In either hypothetical case, you find that you are 30 dollars down from where you thought you were:  either you lost 30 dollars directly, or you lost it indirectly by purchasing a ticket and losing the ticket.  Either way, you lost thirty dollars.  If you are thinking rationally about whether to see the concert, your answer should be the same regardless of how you lost the previous thirty dollars.

The scenario in which the money was lost called an *anchor*.  Behavioral psychologists believe that anchors can affect the way we think about that question, even when the anchor should have no logical bearing on the question itself.

We'll use the data from this survey to investigate the the following Research Question:


  >Who is more likely to purchase a ticket for the rock concert:  a person who finds he/she has lost thirty dollars, or a person who finds he/she has lost a thirty dollar ticket?
  
The data
Study conducted in November 2001 by students in MAT 111. Subjects were 267 Georgetown College students. Not all subjects got the same survey form.  Apparently this data frame had two additional responces added which were **conc.situation** and **conc.decision** and these are the variables of interest in this exercise

Format

A data frame with 268 observations on the following 8 variables.

* **def.race**
Suggested race of the defendant in the survey form.

* **vic.race**
Suggested race of the victim in the survey form.

* **conc.situation**
Scenario described in the in the "rock concert" question on the survey form.

* **sentence**
Sentence, in years, reccommended for the defendant.

* **conc.decision**
Whether or not the subject chose to buy a ticket (or buy another ticket).

* **year**
Class rank of the subject.

* **sex**
a factor with level Sex of the survey participant.

* **major**
possible values: humanities math.sci pre.prof social.sci Type of major the subject intends.

Details

# Methods

<Methods go here.  Perform your variable analysis.  On the basis of that analysis, choose at least one graphical descriptive method, at least one numerical descriptive method, and an appropriate inferential method involving confidence intervals to investigate the Research Question.  NOTE:  Your inferential method should be one that can produce a confidence interval for a parameter of interest.  Suggestion:  In this section you should identify and define the population parameter that you are interested in.>

# Results

<Put some introductory sentence in this section.>

## Descriptive Results

We begin by forming a table between our settings and the decisions made by our subjects.

<Put some introductory sentence in this section.>

### Graphical Descriptive Results
We first look at a barplot of follow on action 

```{r}
contable<-table(attitudes$conc.situation,attitudes$conc.decision)
barchart(contable,
	layout = c(1,2),
	auto.key = list(
		space = "top",
		title = "conc.decision",
		cex.title = 1,
		columns = 1),
	horizontal = FALSE,
	main = "Buy another ticket?")
```

This seems to show losing your ticket is about 50/50 to buying another ... but losing the money makes it a lot more unlikely that you would buy a new one.
<Introduce your graph to the reader, and then make the graph.  After the graph, include a discussion of the graph, pointing out specific features of the graph that help you answer your research question.>

### Numerical Descriptive Results

Lets consider comparing row percents

```{r}
rowPerc(contable)
```

Looking at columes of our table of row percentages we see 80 percent of those who lost tickets, buy again. However only half of those who lost money buy again. This appears to be a noteworthy difference.
<Introduce your numerical descriptive results, and then produce them.  Afterwards, include a discussion of the results, pointing out how specific numbers in the results help you to answer the research question.>

## Inferential Results

In order to quantify the differences observed we will run a chi squared test on the contingency table above.

This is an inferential test done by computing the chi squared static from our sample.

$H_0:$ There is no difference in decision **to buy** based on whether you lost money or a ticket   
$H_a:$ The decision **to buy* is related to whether or not you lost a ticket or money   
We will be interpreting the results by looking at the probablilty of having a chi squared statistic as large as or larger than the value we compute.  This is what the computed p-value represents.  
```{r}
chisqtestGC(contable,graph=TRUE)

```

This pretty well sums it up. P-value extremely low.  Our sample statistic was **26.8**  and the value expected if $H_0:$ is true is right around one.   Very strong evidence against the null hypothesis.



# Discussion and Conclusion

<Draw a general conclusion here, if possible.  If you can't draw a conclusion, say why not. Many people choose to address the Safety Check in this section.  This section is also a good place to suggest new Research Questions suggested by the results of the current study.>